---
title: "Sources and Glossary"
author: "Matt Tabet"
---

This is *yet* another page that can be linked! You are not required to include more than just *index.qmd*, but you are welcome to make content on different pages, if desired. This may help with managing GitHub commits, as well. 

## A note {#sec-note} 

Notice that you can also modify the YAML headings at the top of each page as I did in the [first subpage](another-page.qmd) or completely delete them, as I did here.

```{r}
#| label: setup
#| include: false

# Improve digit display and NA display for kable tables
options(scipen = 1, knitr.kable.NA = "")

# Load necessary packages
library(tidyverse)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyverse)
library(knitr)
library(igraph)
library(tidygraph)
library(ggraph)
library(mdsr)
library(broom)





glossary_data <- data.frame(
  Abbreviation = c("PTS", "FG", "2P", "3P", "FT", "FGA", "2PA", "3PA", "FGX", "AST", "TRB", "STL", "BLK", "TOV", "G", "MP",
                   "FG%", "2P%", "3P%", "FT%", "PPG", "MPG", "APG", "RPG", "SPG", "FTA", "DRB", "ORB", "PF", "TRIP"),
  Name = c("Points", "Field Goals Made", "2 Point Field Goals", "3 Point Field Goals", "Free Throws Made", "Field Goals Attempted",
           "2 Pointers Attempted", "Three Pointers Attempted", "Field Goals Missed", "Assists", "Total Rebounds", "Steals",
           "Blocks", "Turnovers", "Games Played", "Minutes Played", "Field Goal Percentage", "2 Point Percentage", "Three Point Percentage",
           "Free Throw Percentage", "Points Per Game", "Minutes Per Game", "Assists Per Game", "Rebounds Per Game",
           "Steals Per Game", "Free Throws Attempted", "Defensive Rebounds", "Offensive Rebounds", "Personal Fouls", "Triple Doubles"),
  Definition = c("Points scored by adding all made 1s, 2s, and 3s.",
                 "Total made field goals (excluding free throws).",
                 "2-point field goals made.", "3-point field goals made.", "Free throws made.",
                 "Total field goals attempted.", "2-point shots attempted.", "3-point shots attempted.",
                 "Missed field goals.", "Passes leading to made baskets.",
                 "Total offensive and defensive rebounds.", "Steals (excluding deflections).",
                 "Blocks on field goal attempts.", "Turnovers committed.",
                 "Games played.", "Total minutes played.",
                 "FGM / FGA.", "2PM / 2PA.", "3PM / 3PA.", "FTM / FTA.",
                 "Points per game.", "Minutes per game.", "Assists per game.",
                 "Rebounds per game.", "Steals per game.", "Free throws attempted.",
                 "Defensive rebounds.", "Offensive rebounds.", "Personal fouls.",
                 "Double digits in 3 stats in one game.")
)

```

# Glossary

```{r}
#| echo: false
#| warning: false

glossary_data |>
  kable()

```
# Sourcing

We pulled our data from multiple sources off of the internet. For the shot chart, the data was pulled from from an API on NBA.com. For Lebron's general stats and the map of his scoring, we used basketballreference.com and realgm.com. realgm.com was also used to collect data for the teammates of LeBron for both the clustering and network visuals.


